---
title: "R coding Exercise"
output: 
  html_document:
    toc: FALSE
---

#### Load packages  

```{r}
library(tidyverse)
library(ggplot2)
library(dslabs)
library(dplyr)
```
#Look at help file for gapminder data.
```{r}
help("gapminder") 
```

#### Get an overview of the data structure of gapminder  

```{r}
str(gapminder)
```

#### Get a data summary of gapminder.

```{r}
summary(gapminder)
```

#### Determine gapminder's object type  

```{r}
class(gapminder)
```
#### Create new object called "africadata"

```{r}
africadata <- subset(gapminder, continent == "Africa") 
```

#### Get an overview of the data structure of "africadata"  

```{r}
str(africadata)
```

#### Get a data summary of "africadata"  

```{r}
summary(africadata)
```
#### Create new variable in africadata that only contains infant mortality and life expectancy
```{r}
africa_infant <- africadata[c("infant_mortality","life_expectancy")]
str(africa_infant)
summary(africa_infant)
```
#### Create new variable in africadata that only contains population and life expectancy

```{r}
africa_population <- africadata[c("population","life_expectancy")]
str(africa_population)
summary(africa_population)  
```
#### Plot life expectancy as a function of infant mortality.

```{r}
africa_infant_plot <- ggplot(africa_infant, aes(x = log10(infant_mortality), y = life_expectancy)) + geom_point() + ggtitle("Life Expectancy vs. Infant Mortality in Africa") + xlab("Infant Mortality") + ylab("Life Expectancy")
print(africa_infant_plot)
```
#### Plot life expectancy as a function of population.  

```{r}
africa_population_plot <- ggplot(africa_population, aes(x = log10(population), y = life_expectancy)) + geom_point() + ggtitle("Population vs. Infant Mortality in Africa") + xlab("Population") + ylab("Life Expectancy")
print(africa_population_plot)
```












# Filter only records  for year = 2000
```{r}
africadata_2000 <- africadata %>%
  filter(year == 2000)
```


# Cross-checking

```{r}
str(africadata_2000)
```

#Fit-1: Life expectancy as the outcome, and infant mortality as the predictor

```{r}
fit1 <- lm(africadata_2000$life_expectancy ~ africadata_2000$infant_mortality)
summary(fit1)
```

#Fit-2: Life expectancy as the outcome, and population size as the predictor

```{r}
fit2 <- lm(africadata_2000$life_expectancy ~ africadata_2000$population)
summary(fit2)
```

This is extra coding from Dawson Dobash

```{r more figures and tables}
#Scatter plot between life expectancy and gdp. It shows that most countries with high gdp have high life expectancy
africadata_2000 %>% ggplot(aes(x=life_expectancy, y=gdp)) + geom_point()

#Looking at countries with a life expectancy over 60 and displaying the countries in a descending life expectancy order
africadata_2000 %>% filter(life_expectancy > 60) %>% select(life_expectancy,country, region) %>% arrange(desc(life_expectancy))

#Linear model of gdp predicting life expectancy.
fit3 = lm(life_expectancy ~ gdp, data=africadata_2000)

#Summary of the linear model fit3. It shows at a p-value of 0.01, life expectancy is significant in predicting life expectancy
summary(fit3)

#Displaying the mean life expectancy for each region within the year 2000
africadata_2000 %>% select(region, life_expectancy) %>% group_by(region) %>% summarize(average = mean(life_expectancy))
```
